---
title: "R4DS"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
mpg
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

3.3.1
1.
Not blue because colour is inside aes
2.
categorical: manufacturer etc.
continuous: year, displ, hwy etc.
3.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = year))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = manufacturer))
```

continuous var are same colour different shade, vs categoricals which are diff colour
4.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class,shape = class))
```
Doesn't work too well. lol
5.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,  stroke = 5))
```

Doesnt work with all shapes
6.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(colour = displ < 5))
```
Throws error

3.5.1 Ex
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

1.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ hwy, nrow = 2)
```
It looks awful and unhelpful.

2.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

3.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
Plot hwy by displ for each drv and cyl category
The . removes faceting either col or rows depending on which side of the ~
4. 
