---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(forcats)
```
```{r}
x <- c("Jan","Mar","Feb")

month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

y <- factor(x, levels = month_levels)

sort(y)
```

```{r}
gss_cat

gss_cat %>% 
  ggplot(aes(race))+geom_bar()
```

```{r}
gss_cat %>% 
  ggplot(aes(rincome))+geom_bar()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
I could improve this chart by fixing the orientation of the x axis labels

```{r}
gss_cat %>% count(relig)
```
Most common religion is protestant
```{r}
gss_cat %>% count(partyid)
```
Most common political affiliation is not strong democrat
```{r}
gss_cat %>% 
  select(relig,denom) %>% 
  filter(relig != "Protestant")

```
denom applies to protestant, in the table u can see that for any non protestant relig, denom is not applicable


```{r}
gss_cat %>%
  count(relig, denom) %>%
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```

15.4

```{r}
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )
relig_summary %>%
  mutate(relig = fct_reorder(relig, tvhours)) %>%
  ggplot(aes(tvhours, relig)) +
    geom_point()
```

```{r}
by_age <- gss_cat %>%
  filter(!is.na(age)) %>%
  count(age, marital) %>%
  group_by(age) %>%
  mutate(prop = n / sum(n))

ggplot(by_age, aes(age, prop, colour = marital)) +
  geom_line(na.rm = TRUE)

ggplot(by_age, aes(age, prop, colour = fct_reorder2(marital, age, prop))) +
  geom_line() +
  labs(colour = "marital")
```

There are some suspiciously high numbers in tvhours. Is the mean a good summary?
```{r}
gss_cat %>% 
  ggplot(aes(tvhours))+geom_bar()
```
mean isn't the worst but i think median would be better

For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.
```{r}
gss_cat
```
martial - arbit
race - arbit
rincome - principled
partyid - arbit
relig  -arbit
denom  -arbit
tvhours - principled

Why did moving "Not applicable" to the front of the levels move it to the bottom of the plot?
```{r}

```

15.5
How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?
```{r}
gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  count(year,partyid) %>% 
  group_by(year) %>% 
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(x=year, y=p, color=fct_reorder2(partyid,year,p)))+
  geom_point()+
  geom_line()
```
```{r}
60/(684+1152+921)
```

How could you collapse rincome into a small set of categories?

Group all the small categories together, group the non responders and group the wealthy together as these factors are numerous but have low counts

