---
title: "R Notebook"
output: html_notebook
---

Iteration methods
1.imperative
2.functional

```{r}
library(tidyverse)
```

simple loop example
```{r}
df <- tibble(a = rnorm(10),
       b = rnorm(10),
       c = rnorm(10),
       d = rnorm(10))

```

```{r}
output <- vector("double", ncol(df))  # 1. output, make empty 'space' for the output to fill later
for (i in seq_along(df)) {            # 2. sequence, iterate along this vector ...
  output[[i]] <- median(df[[i]])      # 3. body, stick the output in the empty data structure we created using func 
}
output # read out
```

21.2.1 Exercises
Write for loops to:

Compute the mean of every column in mtcars.
```{r}
mtcars
```
```{r}
mtcarsMean <- vector("double", length = ncol(mtcars))
for (i in seq_along(mtcars)) {
  mtcarsMean[[i]] <- mean(mtcars[[i]])
}
mtcarsMean
```

Determine the type of each column in nycflights13::flights.
```{r}
library('nycflights13')

nycfcolType <- vector("character", length= ncol(nycflights13::flights))
for (i in names(nycflights13::flights)) {
  nycfcolType[[i]] <- typeof(nycflights13::flights[[i]])
}

nycfcolType
```

Compute the number of unique values in each column of iris.
```{r}
install.packages("iris")
library("iris")
```
```{r}
iris
```
```{r}
uniq_iris <- vector("double", ncol(iris))
for (i in seq_along(iris)) {
  uniq_iris[i] <- length(unique(iris[[i]]))
}

uniq_iris
```

Generate 10 random normals for each of  
??
=
???
10
 ,  
0
 ,  
10
 , and  
100
 .
Think about the output, sequence, and body before you start writing the loop.

```{r}
mew <- c(-10,0,10,100)

norms <- vector("list", length(mew))
for (i in seq_along(norms)) {
  norms[[i]] <- rnorm(10, mean = mew[i])
}

norms
```

Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:

out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}
```{r}
letters

str_c(letters, sep = " ", collapse = "")
```

x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))
```{r}
sd(rnorm(100))
```

x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}

Combine your function writing and for loop skills:

Write a for loop that prints() the lyrics to the children's song "Alice the camel".

Convert the nursery rhyme "ten in the bed" to a function. Generalise it to any number of people in any sleeping structure.

Convert the song "99 bottles of beer on the wall" to a function. Generalise to any number of any vessel containing any liquid on any surface.

It's common to see for loops that don't preallocate the output and instead increase the length of a vector at each step:

```{r}
x <- c(10,2,4,1)
output <- vector("integer", 0)
for (i in seq_along(x)) {
  output <- c(output, lengths(x[[i]]))
}
output
```

How does this affect performance? Design and execute an experiment

makes it slower.

##################################

There are four variations on the basic theme of the for loop:

Modifying an existing object, instead of creating a new object.
Looping over names or values, instead of indices.
Handling outputs of unknown length.
Handling sequences of unknown length.

####################################

There are three basic ways to loop over a vector. So far I've shown you the most general: looping over the numeric indices with for (i in seq_along(xs)), and extracting the value with x[[i]]. There are two other forms:

Loop over the elements: for (x in xs). This is most useful if you only care about side-effects, like plotting or saving a file, because it's difficult to save the output efficiently.

Loop over the names: for (nm in names(xs)). This gives you name, which you can use to access the value with x[[nm]]. This is useful if you want to use the name in a plot title or a file name. If you're creating named output, make sure to name the results vector like so:

results <- vector("list", length(x))
names(results) <- names(x)

Iteration over the numeric indices is the most general form, because given the position you can extract both the name and the value:

for (i in seq_along(x)) {
  name <- names(x)[[i]]
  value <- x[[i]]
}

###################################

Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, files <- dir("data/", pattern = "\\.csv$", full.names = TRUE), and now want to read each one with read_csv(). Write the for loop that will load them into a single data frame.
```{r}
dfcsv <- vector("character", length(files))
for (i in files) {
  dfcsv[[i]] <- read_csv(files[[i]])
}

dfcsv <- bind_rows(dfcsv)
```

What happens if you use for (nm in names(x)) and x has no names? What if only some of the elements are named? What if the names are not unique?
```{r}
x <- c(1:10)
for (nm in names(x)) {
  print(nm)  
}
y <- c(a = 1, a = 2, b = 3, 6)
for (nm in names(y)) {
  print(nm)  
}
```
wont do anything, will print out duplicates

Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, show_mean(iris) would print:

show_mean(iris)
```{r}
show_mean <- function(x) {
  meanvector <- vector("double",length = ncol(x))
  for (i in seq_along(x)) {
    meanvector[[i]] <- mean(x[[i]])
  }
}

show_mean(iris)

```

```{r}
show_mean <- function(df, digits = 2) {
  # Get max length of all variable names in the dataset
  maxstr <- max(str_length(names(df)))
  for (nm in names(df)) {
    if (is.numeric(df[[nm]])) {
      cat(
        str_c(str_pad(str_c(nm, ":"), maxstr + 1L, side = "right"),
          format(mean(df[[nm]]), digits = digits, nsmall = digits),
          sep = " "
        ),
        "\n"
      )
    }
  }
}
show_mean(iris)
```

(Extra challenge: what function did I use to make sure that the numbers lined up nicely, even though the variable names had different lengths?)

What does this code do? How does it work?
```{r}
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}

trans
```
take disp column and multiply it by .01638

```{r}
a <- tibble(a = rnorm(10), b = rnorm(10))
a
```
```{r}
apply(a, 2, sum)
```

map functions (+purrr)
design pattern: loop over a vector, doing something to each element and save the result to memory
```{r}
map_dbl(a, mean)
```

```{r}
library(purrr)
```


```{r}
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))
models

```
Write code that uses one of the map functions to:

Compute the mean of every column in mtcars.
```{r}
map_dbl(mtcars, mean)
```

Determine the type of each column in nycflights13::flights.
```{r}
map_chr(nycflights13::flights,typeof)
```

Compute the number of unique values in each column of iris.
```{r}
map_dbl(iris, function(x) length(unique(x)))
```

Generate 10 random normals for each of  
??
=
???
10
 ,  
0
 ,  
10
 , and  
100
 .
```{r}
map_dbl(c(-10,0,10,100), ~ rnorm(n = 10, ))
```


How can you create a single vector that for each column in a data frame indicates whether or not it's a factor?

What happens when you use the map functions on vectors that aren't lists? What does map(1:5, runif) do? Why?

What does map(-2:2, rnorm, n = 5) do? Why? What does map_dbl(-2:2, rnorm, n = 5) do? Why?

Rewrite map(x, function(df) lm(mpg ~ wt, data = df)) to eliminate the anonymous function.


```{r}

```

