---
title: "R4DS"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
mpg
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

3.3.1
1.
Not blue because colour is inside aes
2.
categorical: manufacturer etc.
continuous: year, displ, hwy etc.
3.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = year))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = manufacturer))
```

continuous var are same colour different shade, vs categoricals which are diff colour
4.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class,shape = class))
```
Doesn't work too well. lol
5.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,  stroke = 5))
```

Doesnt work with all shapes
6.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(colour = displ < 5))
```
Throws error

3.5.1 Ex
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

1.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ hwy, nrow = 2)
```
It looks awful and unhelpful.

2.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

3.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
Plot hwy by displ for each drv and cyl category
The . removes faceting either col or rows depending on which side of the ~
4. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

Can see multiple plots, instead of one making it easier to identify trends than all one one plot with diff colours

5.
facet_wrap wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than facet_grid() because most displays are roughly rectangular.

nrow specifies to split into how many rows. scale and shrink also affect size and output

facet_grid forms a matrix of panels defined by row and column faceting variables. It is most useful when you have two discrete variables, and all combinations of the variables exist in the data.

grid has no rowcol options because it depends on the variables being faceted

6.
The variable with more levels should be in the col in a facetgrid because it easier to scale the y axis vs the x axis in terms of numeric value

3.6
1.
```{r}
geom_line()
geom_boxplot()
geom_histogram()
geom_area()
```
2.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
3.
Removes the legend from graph

4.
se shows confidence interval aroun smooth line

5.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

6.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se=FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(mapping = aes(x=displ,y=hwy,group=drv,),se=FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy,color=drv)) + 
  geom_point() + 
  geom_smooth(mapping = aes(x=displ,y=hwy,group=drv,),se=FALSE)
```

3.7.1
1. 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

3.7.1
1. def for stat summ is a pointrange

```{r}
ggplot(data = diamonds) + 
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat='summary',
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
    )
#fun is a function
```

2.
geom_col uses raw num, and req a y value and geom_bar uses a proportionate value for each x. geom_bar uses stat_summary to count each x to get y
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

ggplot(data = diamonds) + 
  geom_col(mapping = aes(x = cut, y = 1))
```

4.
stat_smooth is like geom_smooth, makes confidence intervals given y and x

5.
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```
need group=1 or else it will assume all proportions are the same

3.8.1
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
ggplot(data = diamonds) + 
  geom_bar(alpha=.2,mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

1.
Overlapping from overplotting, we can add some jitter to distinguish the points
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

2. the width and height parameters control jitter amount

3.geom_jitter vs geom_count
the jitter geom adds variance to each point where count geom aggregates overlapping points by increasing point size
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
```

4.
From the differences in the size of the boxes, the default geom is count
```{r}
ggplot(data = mpg,mapping = aes(x = displ, y = hwy, color = class)) + 
  geom_boxplot()
```

3.9.1
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```
1.
```{r}
bar<-ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))

bar+coord_polar()
```
2.
labs() changes the labels in the viz

3.
quickmap is an approximation whereas in map you have more control over parameter values of the function

4.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

coord_fixed fixes the aspect ratio of the axes
the relationship is positive and that for approx increasing cty by 5 will increase hwy by 10.

3.10
the template 

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
































