---
title: "R4DS CH5"
output: html_notebook
---

```{r}
library(tidyverse)
install.packages('nycflights13')
```


1.
```{r}
library(nycflights13)
flights
filter(flights, month == 1,day == 2)
```

Had an arrival delay of two or more hours
```{r}
filter(flights,arr_delay > 2)
```

Flew to Houston (IAH or HOU)
```{r}
filter(flights,dest == "IAH" | dest == "HOU")
```

Were operated by United, American, or Delta
```{r}
unique(flights["carrier"])
```
```{r}
filter(flights, carrier == "DL"| carrier == "UA"| carrier == "AA")
```

Departed in summer (July, August, and September)
```{r}
filter(flights,month >= 7 ,month <=9)
```
Arrived more than two hours late, but didn't leave late
```{r}
filter(flights, arr_delay > 2, dep_delay <= 0)
```

Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 1, dep_delay - arr_delay > 30)
```

Departed between midnight and 6am (inclusive)
```{r}
filter(flights, sched_dep_time <= 600 | sched_dep_time == 2400 )
```

2.
between() returns if values fall between a range
```{r}
filter(flights, between(sched_dep_time, 0, 600))
```

3.
```{r}
filter(flights,is.na(dep_time))
```
Probably cancellations

4.
anything to power 0 equals one so not NA
```{r}
NA^0
```

5.3.1
1.
```{r}
arrange(flights, desc(is.na(dep_time)),dep_time)
```

2.
```{r}
arrange(flights, desc(dep_delay))
```

3.
```{r}
arrange(flights,air_time)
arrange(flights,desc(air_time))
```

5.4.1
1.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)


```

```{r}
select(flights,dep_time:arr_time, arr_delay)
```

```{r}
select(flights, contains("delay"), contains("time"))
```

2.
```{r}
select(flights, arr_time, arr_time)
```

nothing

3.
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights,one_of(vars))
```

one_of matches variable names to those stored in a vector, nice!

4.
```{r}
select(flights,contains("TIME", ignore.case = FALSE))
```

yes b/c its not case sensitive just like sql lol

5.5.2
1.
```{r}
deps <- select(flights, arr_time, dep_time)

mutate(deps, dep_min_since_midnight = ((dep_time%/%100)*60 + dep_time%%100) %% 1440)
```

2.
```{r}
airtime<-select(flights, arr_time, dep_time, air_time)
mutate(airtime , test = arr_time - dep_time)
```

new var is greater than air_time,  might not include time sitting on the tarmac

3.
```{r}
select(flights, dep_time, sched_dep_time, dep_delay)
```

Would expect sched dep time and delay to equal dep time

4.
```{r}
delays<-mutate(flights, dep_delay_rank = min_rank(-dep_delay))

delays <- filter(delays, dep_delay_rank <= 20)

arrange(delays, dep_delay_rank)
```

5.
```{r}
1:3 + 1:10
```
becase the frist vector is too short so it just repeats itself [1,2,3,1,2,3,1,2,3,1] over to match length of the second

5.6.1
1.
```{r}
summarise(flights,delay=mean(dep_delay, na.rm = TRUE))

```

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

2.
```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

```{r}
flights %>%
  select(everything()) %>%  
  summarise_all(funs(sum(is.na(.))))
```

Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
```{r}
#over and under median time, +- 15?
ss_flights <- select(flights, flight, year, month, day, sched_dep_time, dep_delay)
ss_flights

flights %>% 
  group_by(flight) %>% 
  summarize(early15 = sum(arr_delay <= -15, na.rm = TRUE) / n(),
            late15 = sum(arr_delay >= 15, na.rm = TRUE) / n()) %>%
filter(early15 == .5, late15 == .5)

```

A flight is always 10 minutes late.
```{r}
flights %>% 
  group_by(flight) %>% 
  summarise(late10 = sum(arr_delay == 10, na.rm = TRUE)/n()) %>% 
  filter(late10 == 1)
  
```

A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
```{r}
flights %>% 
  group_by(flight) %>% 
  summarize(early30 = sum(arr_delay <= -30, na.rm = TRUE) / n(),
            late30 = sum(arr_delay >= 30, na.rm = TRUE) / n()) %>%
filter(early30 == .5, late30 == .5)
```

99% of the time a flight is on time. 1% of the time it's 2 hours late.
```{r}
flights %>% 
  group_by(flight) %>% 
  summarize(on99 = sum(arr_delay = 0, na.rm = TRUE) / n(),
            late2hr = sum(arr_delay >= 120, na.rm = TRUE) / n()) %>%
filter(on99 == .99, late2hr == .01)
```

Which is more important: arrival delay or departure delay?
For me probably arrival.

Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% count(dest)
```
```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(n = n())
```
```{r}
not_cancelled %>%
  count(tailnum, wt = distance)

```
```{r}
not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(n = sum(wt = distance))
```

Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

The defn is as follows: prob of dep delay being NA given arr delay being NA is a cancelled flight. The most important column should be airtime or  something

```{r}
filter(flights, !is.na(dep_delay), is.na(arr_time)) %>% 
  select(everything())
```

Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
weird <- flights %>% 
  filter(is.na(dep_time),is.na(arr_time)) %>% 
  select(everything())

weird2  <- flights %>%
  group_by(year, month, day) %>%
  summarize(prob_cancelled = sum(is.na(dep_delay)) / n(),
            avg_delay = mean(dep_delay, na.rm = TRUE))

ggplot(data = weird2, mapping = aes(x = prob_cancelled, y= avg_delay))+
  geom_point()+
  geom_smooth()                                                                         

```


Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

Bad airports vs bad carriers ... lol see the avg delay of each airport vs each carrier
```{r}
flights %>% group_by(carrier, dest) %>% summarise(n())
```
```{r}
flights %>% 
  group_by(carrier) %>% 
  summarise(meandelay = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(meandelay))

filter(airlines,carrier == "F9")
```


What does the sort argument to count() do. When might you use it?

sorts output in descending order if TRUE, used bfore doing an arrange action
```{r}
count(flights, year, sort=TRUE)
```

5.7 grouped mutates
2. which tailnum has worst on time
```{r}
flights %>% 
  group_by(tailnum) %>% 
  summarise(ontime=mean(arr_delay)) %>% 
  arrange(desc(ontime)) %>% 
  select(tailnum,ontime)
```

What time of day should you fly if you want to avoid delays as much as possible?
```{r}
names(flights)
```
```{r}
flights %>% 
  group_by(time_hour) %>% 
  summarise(ontime2=mean(arr_delay)) %>% 
  arrange(desc(ontime2)) %>% 
  select(time_hour,ontime2)
```

For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.

```{r}
flights %>% 
  group_by(dest) %>% 
  summarise(destdelay = sum(arr_delay,na.rm = TRUE)) %>% 
  arrange(desc(destdelay)) %>% 
  select(dest,destdelay)
```

Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag(), explore how the delay of a flight is related to the delay of the immediately preceding flight.
```{r}
lagged_delays <- flights %>%
  arrange(origin, year, month, day, dep_time) %>%
  group_by(origin) %>%
  mutate(dep_delay_lag = lag(dep_delay)) %>%
  filter(!is.na(dep_delay), !is.na(dep_delay_lag))

lagged_delays %>%
  summarise(delay_diff = mean(dep_delay - dep_delay_lag), na.rm = TRUE)

lagged_delays %>%
  group_by(dep_delay_lag) %>%
  summarise(dep_delay_mean = mean(dep_delay)) %>%
  ggplot(aes(y = dep_delay_mean, x = dep_delay_lag)) +
  geom_point() +
  geom_smooth() +
  labs(y = "Departure Delay", x = "Previous Departure Delay")
```

Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.
```{r}
twocarrier <- select(flights,carrier,dest)

twocarrier %>% 
  group_by(dest) %>% 
  mutate(n_carrier = n_distinct(carrier)) %>%
  filter(n_carrier >= 2) %>% 
  distinct()
```

For each plane, count the number of flights before the first delay of greater than 1 hour

```{r}
flights %>%
  arrange(tailnum, year, month, day) %>%
  group_by(tailnum) %>%
  mutate(delay_gt1hr = dep_delay > 60) %>%
  mutate(before_delay = cumsum(delay_gt1hr)) %>%
  filter(before_delay < 1) %>%
  count(sort = TRUE)
```






































